---
title: "VDJdb validation analytics"
author: "M.S."
date: "2025-07-16"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggalluvial)
library(ggh4x)
library(patchwork)
library(stringr)
library(showtext)
showtext_auto(T)
font_add("Helvetica light", "assets/HelveticaLight.ttf")
font_add("Helvetica", "assets/HelveticaNeueRoman.ttf")
library(png)
library(grid)
library(reshape2)
```

### Load data

Load VDJdb table - here we use the version that matches the date of Messemaker et al. preprint release.

```{r}
data <- read_tsv("dump/vdjdb-2025-02-21/vdjdb_full.txt") |>
  filter(species == "HomoSapiens") |>
  select(v.alpha, j.alpha, cdr3.alpha, 
         v.beta, j.beta, cdr3.beta,
         antigen.epitope, mhc.a, mhc.b,
         method.identification,
         method.singlecell,
         reference.id) |>
  unique()
glimpse(data)
```

```{r}
data |>
  filter(!is.na(cdr3.alpha)) |>
  nrow()
data |>
  filter(!is.na(cdr3.beta)) |>
  nrow()
data |>
  filter(!is.na(cdr3.alpha), !is.na(cdr3.beta)) |>
  nrow()
length(unique(data$antigen.epitope))
```

Also load TCRNET motifs

```{r}
data.motifs <- read_tsv("dump/vdjdb-2025-02-21/cluster_members.txt") |>
  filter(species == "HomoSapiens")
glimpse(data.motifs)
```

### Summary of antigen specificity database biases

Compare imbalance in terms of uneven HLA/epitope coverage. Also check how this imbalance drops when going from records to motifs. Note that we can compute and compare Gini indices which would be more statistically sound, however, for the sake of simplicity for the reader we just explore the case of top 1 by coverage vs median.

By MHC, we take only first chain focusing on MHCI, there are obviously far few MHCII

```{r}
data.s.mhc <- data |>
  mutate(mhc.a = str_split_fixed(mhc.a, fixed(":"), 2)[,1]) |>
  group_by(mhc.a) |>
  summarise(records = n(), publications = length(unique(reference.id))) |>
  ungroup() |>
  mutate(fraction.records = records / sum(records) * 100,
         fraction.publications = publications / sum(publications) * 100) |>
  arrange(-records)

data.s.mhc |>
  head()

data.s.mhc |>
  summarise(records.m = median(records),
            publications.m = median(publications))

data.motifs.s.mhc <- data.motifs |>
  mutate(mhc.a = str_split_fixed(mhc.a, fixed(":"), 2)[,1]) |>
  group_by(gene, mhc.a) |>
  summarise(records = n(), motifs = length(unique(cid))) |>
  ungroup() |>
  mutate(fraction.records = records / sum(records) * 100,
         fraction.motifs = motifs / sum(motifs) * 100) |>
  arrange(-records)

data.motifs.s.mhc |>
  head()

data.motifs.s.mhc |>
  group_by(gene) |>
  summarise(records.m = median(records),
            motifs.m = median(motifs))
```

By epitope

```{r}
data.s.ag <- data |>
  group_by(antigen.epitope) |>
  summarise(records = n(), publications = length(unique(reference.id))) |>
  ungroup() |>
  mutate(fraction.records = records / sum(records) * 100,
         fraction.publications = publications / sum(publications) * 100) |>
  arrange(-records)

data.s.ag |>
  head()

data.s.ag |>
  summarise(records.m = median(records),
            publications.m = median(publications))

data.motifs.s.ag <- data.motifs |>
  group_by(gene, antigen.epitope) |>
  summarise(records = n(), motifs = length(unique(cid))) |>
  ungroup() |>
  mutate(fraction.records = records / sum(records) * 100,
         fraction.motifs = motifs / sum(motifs) * 100) |>
  arrange(-records)

data.motifs.s.ag |>
  head()

data.motifs.s.ag |>
  group_by(gene) |>
  summarise(records.m = median(records),
            motifs.m = median(motifs))
```

### Overview of VDJdb confidence scoring features

By assay type

```{r}
method_variants <- summary(as.factor(data$method.identification))
tibble(method = names(method_variants), records = method_variants) |>
  arrange(-records)
```

```{r}
summary(as.factor(data$method.singlecell))
```

```{r}
data.assay <- data |>
  mutate(assay = case_when(method.singlecell == "yes" & 
                             grepl("sort", method.identification) ~ '3.scRNA-seq',
                           (is.na(method.singlecell) | method.singlecell != "yes") & 
                             grepl("sort", method.identification) ~ '2.MHC-multimer FACS',
                           (!is.na(method.identification) & 
                             !grepl("sort", method.identification)) ~ '1.immunoassay',
                           T ~ 'other'),
         chain = case_when(is.na(cdr3.beta) ~ "TRA only",
                           is.na(cdr3.alpha) ~ "TRB only",
                           !is.na(cdr3.beta) & !is.na(cdr3.alpha) ~ "paired"))

data.assay.s <- data.assay |>
  group_by(chain, assay) |>
  summarise(records = n(), studies = length(unique(reference.id)))
```

```{r}
coef = 200
brks = c(0, 10000, 20000, 30000, 40000, 50000, 60000)
p1 <- ggplot(data.assay.s, aes(x = assay,
                         fill = chain)) +
  geom_bar(aes(y = records), 
           stat="identity", 
           color = "black", linewidth = 0.3) +
  geom_bar(aes(y = -studies*coef), 
           stat="identity", 
           color = "black", linewidth = 0.3) +
  geom_hline(yintercept = 0, 
             alpha = 0.5,
             linewidth = 0.3) +
  scale_y_continuous(
    # Features of the first axis
    name = "records",
    breaks = brks,
    limits = c(-30000, 45000),
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*-coef, name="studies",
                        breaks = brks * coef,
                        labels = brks / coef)
  ) +
  scale_fill_manual("",
                    breaks = c("TRA only",
                               "TRB only", 
                               "paired"),
                    values = c("#abd9e9", "#74add1", "#fdae61")) +
  scale_x_discrete("assay", labels = c("1.IA", "2.FACS", "3.sc", "other")) +
  theme_classic() +
  theme(
    text=element_text(size=12, family="Helvetica light"),
    axis.line = element_line(color='black', linewidth = 0.3),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(hjust=0.8),
    legend.background = element_blank(),
    legend.position = c(0.30, 0.97),
    legend.text = element_text(size=8),
    legend.title = element_text(size=10)#,
    #axis.text.x = element_text(angle=90, vjust=.5, hjust=1)
    )
p1
```

By number of studies

```{r}
data.slim <- read_tsv("dump/vdjdb-2025-02-21/vdjdb.slim.txt")
glimpse(data.slim)

data.score <- data.slim |>
  separate_rows(reference.id, sep = ",") |>
  group_by(gene, v.segm, j.segm, cdr3) |>
  mutate(studies = length(unique(reference.id))) |>
  mutate(studies = ifelse(studies > 2, "3+", as.character(studies))) |>
  group_by(vdjdb.score, studies) |>
  summarise(clonotypes = n())
```

```{r}
p1a <- ggplot(data.score |>
         mutate(vdjdb.score = case_when(
           vdjdb.score == 1 ~ "Low",
           vdjdb.score == 2 ~ "Med",
           vdjdb.score == 3 ~ "High",
           T ~ "N/A"
         ) |> factor(levels = c("N/A", "Low", "Med", "High"))), 
       aes(x = studies, y = clonotypes, fill = vdjdb.score)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", linewidth = 0.3) +
  #facet_wrap(~ vdjdb.score, ncol = 1, strip.position = "left", dir = "v") + 
  scale_y_log10(breaks = c(1, 10, 100, 1000, 10000, 50000)) +
  scale_fill_manual("VDJdb\nconf.",
                    breaks = c("N/A",
                               "Low", 
                               "Med",
                               "High"),
                    values = c("#f1b6da", "#f7f7f7", "#b8e186", "#7fbc41")) +
  theme_classic() +
  theme(
    text = element_text(size=12,  family="Helvetica light"),
    axis.line = element_line(color='black', linewidth = 0.3),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.title.y = element_text(hjust=0.8),
    legend.background = element_blank(),
    legend.position = c(0.8, 0.9),
    legend.text = element_text(size=8),
    legend.title = element_text(size=10)#,
    #axis.text.x = element_text(angle=90, vjust=.5, hjust=1)
    )

p1a
```

More detailed view of studies per clonotype by assay

```{r}
data.epi <- rbind(data.assay |>
                        filter(!is.na(cdr3.beta)) |>
                        group_by(cdr3.beta) |>
                        mutate(
                          chain = "TRB",
                          epitopes = length(unique(antigen.epitope))),
                      data.assay |>
                        filter(!is.na(cdr3.alpha)) |>
                        group_by(cdr3.alpha) |>
                        mutate(
                          chain = "TRA",
                          epitopes = length(unique(antigen.epitope))))|>
  group_by(assay, chain, epitopes) |>
  summarise(clonotypes = length(unique(paste(v.alpha, cdr3.alpha, j.alpha,
                                          v.beta, cdr3.beta, j.beta))))

data.studies <- rbind(data.assay |>
                        filter(!is.na(cdr3.beta)) |>
                        group_by(cdr3.beta) |>
                        mutate(
                          chain = "TRB",
                          studies = length(unique(reference.id))),
                      data.assay |>
                        filter(!is.na(cdr3.alpha)) |>
                        group_by(cdr3.alpha) |>
                        mutate(
                          chain = "TRA",
                          studies = length(unique(reference.id))))|>
  group_by(assay, chain, studies) |>
  summarise(clonotypes = length(unique(paste(v.alpha, cdr3.alpha, j.alpha,
                                          v.beta, cdr3.beta, j.beta))))
```

```{r}
p2 <- ggplot(data.studies |>
               filter(assay != "other"), 
             aes(x = studies, y = clonotypes, color = chain)) +
  geom_path() +
  geom_point() +
  scale_y_log10() +
  scale_x_continuous(breaks = c(1, 5, 10, 15)) +
  scale_color_manual("",
                    breaks = c("TRA", "TRB"),
                    values = c("#abd9e9", "#74add1")) +
  facet_wrap(~assay) +
  theme_classic() +
  theme(
    text=element_text(size=12,  family="Helvetica light"),
    axis.line = element_line(color='black', linewidth = 0.3),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.background = element_blank(),
    strip.background = element_blank(),
    legend.position = c(0.90, 0.85),
    legend.text = element_text(size=8),
    legend.title = element_text(size=10))
p2
```

```{r}
p3 <- ggplot(data.epi |>
               filter(assay != "other"), 
             aes(x = epitopes, y = clonotypes, color = chain)) +
  geom_path() +
  geom_point() +
  scale_y_log10() +
  scale_x_continuous(breaks = c(1, 5, 10, 15)) +
  scale_color_manual(guide = F,
                    breaks = c("TRA", "TRB"),
                    values = c("#abd9e9", "#74add1")) +
  facet_wrap(~assay) +
  theme_classic() +
  theme(
    text=element_text(size=12,  family="Helvetica light"),
    axis.line = element_line(color='black', linewidth = 0.3),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(size = 0.5, fill = 'white', color = "white"),
    strip.text = element_text(colour = 'white'))
p3
```

Number of supporting studies for records that fall into motifs 

```{r}
data.motifs.1 <- rbind(data.assay |>
                      filter(!is.na(cdr3.beta)) |>
                      mutate(in.motif = cdr3.beta %in% data.motifs$cdr3aa) |>
                      group_by(cdr3.beta) |>
                      mutate(
                        chain = "TRB",
                        epitopes = length(unique(antigen.epitope)),
                        studies = length(unique(reference.id))),
                    data.assay |>
                      filter(!is.na(cdr3.alpha)) |>
                      mutate(in.motif = cdr3.alpha %in% data.motifs$cdr3aa) |>
                      group_by(cdr3.alpha) |>
                      mutate(
                        chain = "TRA",
                        epitopes = length(unique(antigen.epitope)),
                        studies = length(unique(reference.id)))) |>
  group_by(chain, in.motif, epitopes, studies) |>
  summarise(clonotypes = length(unique(paste(v.alpha, cdr3.alpha, j.alpha,
                                          v.beta, cdr3.beta, j.beta)))) |>
  group_by(chain, in.motif) |>
  mutate(total_clonotypes = sum(clonotypes)) |>
  ungroup()
data.motifs.1
```

```{r}
p4 <- ggplot(data.motifs.1 |>
               mutate(in.motif = ifelse(in.motif, "TCRNET+", "orphan")), 
             aes(x = epitopes, y = studies)) +
  geom_tile(color = "grey25", size = 0.3, 
           #binwidth = c(1, 1),
           aes(fill = clonotypes / total_clonotypes * 100)) +
  scale_fill_gradient2("%clono-\n types",
                       trans = "log10",
                       mid = "#ffffbf", low = "#4575b4", high = "#f46d43",
                       breaks = c(0.01, 0.1, 1, 10, 50),
                       midpoint = 0.1) +
  scale_x_continuous(breaks = c(1, 5, 10, 15)) +
  scale_y_continuous(breaks = c(1, 5, 10, 15)) +
  #scale_y_continuous(limits = c(0, 17), breaks = c(1, 5, 10, 15)) +
  #scale_y_continuous(limits = c(0, 17), breaks = c(1, 5, 10, 15), expand = expansion(0, 1)) +
  #facet_grid(~ chain + in.motif) +
  facet_nested( ~ chain + in.motif, nest_line = T) +
  theme_classic() +
  theme(
    #aspect.ratio = 1,
    text = element_text(size=12,  family="Helvetica light"),
    axis.line = element_line(color = 'black', linewidth = 0.3),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    legend.background = element_blank(),
    legend.position = c(0.97, 0.6),
    legend.text = element_text(size=8),
    legend.title = element_text(size=10)
    #strip.background = element_rect(size = 0.5)
    )

p4
```

### Matchmakers data analysis

First, just estimate using numbers from Messemaker et al. We don't know outcome for each record,
just use reported validation rates as priors.

```{r}
data.slim |>
  filter(antigen.epitope %in% c("YLQPRTFLL", "GLCTLVAML"),
         complex.id != 0) |>
  mutate(in.motif = ifelse(cdr3 %in% data.motifs$cdr3aa,
                           "TCRNET+", "orphan")) |>
  mutate(vdjdb.score = case_when(
    vdjdb.score == 1 ~ "Lo",
    vdjdb.score == 2 ~ "Med",
    vdjdb.score == 3 ~ "Hi",
    T ~ "N/A"
  ) |> factor(levels = c("N/A", "Lo", "Med", "Hi"))) |>
  separate_rows(reference.id, sep = ",") |>
  group_by(gene, v.segm, j.segm, cdr3) |>
  mutate(studies = length(unique(reference.id))) |>
  mutate(studies = ifelse(studies > 2, "3+", as.character(studies))) |>
  group_by(antigen.epitope, vdjdb.score, studies, in.motif) |>
  summarise(clonotypes = length(unique(paste(v.segm, j.segm, cdr3)))) -> data.alluvial

# Matchmakers estimate
tibble(antigen.epitope = c("YLQPRTFLL", "GLCTLVAML", "YLQPRTFLL", "GLCTLVAML"),
       matchmakers = c("MM+", "MM+", "MM-", "MM-"),
       mm_frac_ae = c(0.406, 0.694, 1 - 0.406, 1 - 0.694)) |>
  full_join(tibble(vdjdb.score = c("N/A", "Lo", "Med", "Hi", "N/A", "Lo", "Med", "Hi"),
                   matchmakers = c("MM+", "MM+", "MM+", "MM+", "MM-", "MM-", "MM-", "MM-"),
                   mm_frac_vs = c(0.65, 0.45, 0.8, 1.0, 1.0 - 0.65, 1.0 - 0.45, 1.0 - 0.8, 1.0 - 1.0)) |>
              group_by(matchmakers) |>
              mutate(mm_frac_vs = mm_frac_vs / sum(mm_frac_vs)),
            by = "matchmakers", relationship = "many-to-many") -> data.mm

data.alluvial.mm <- full_join(data.alluvial, 
                              data.mm, 
                              by = c("antigen.epitope", "vdjdb.score"),
                              relationship = "many-to-many") |>
  mutate(frac = clonotypes * mm_frac_ae * mm_frac_vs)
```

```{r}
img.vdjdb <- rasterGrob(readPNG("assets/vdjdb.png"), interpolate=T)
img.mm <- rasterGrob(readPNG("assets/mm.png"), interpolate=T)

p5 <- ggplot(data = data.alluvial.mm |>
         filter(vdjdb.score != "N/A") |>
         group_by(antigen.epitope, studies,
                  vdjdb.score, in.motif, matchmakers) |>
         summarise(frac = sum(frac)),
       aes(axis1 = studies,
           axis2 = antigen.epitope,
           axis4 = vdjdb.score,
           axis5 = in.motif,
           axis6 = matchmakers,
           y = frac)) +
  scale_x_discrete("",
                   limits = c("Studies", "Antigen", "VDJdb score", "Motif", "Validation"), 
                   expand = c(.5, .05),
                   position = "bottom") +
  ylab("") +
  geom_alluvium(aes(fill = matchmakers), color = NA) +
  scale_fill_manual(guide = F, values = c("#de77ae", "#7fbc41")) +
  geom_stratum(color = "black", fill = "white", size = 0.2) +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            angle = 90, family="Helvetica light",
            cex = 3) +
  annotation_custom(img.vdjdb, xmin=-1.0, xmax=0.75, ymin=-Inf, ymax=Inf) +
  annotation_custom(img.mm, xmin=5.3, xmax=7, ymin=-Inf, ymax=Inf) +
  theme_classic() +
  theme(text = element_text(size=12, family="Helvetica light"),
        axis.line.y = element_blank(),
        axis.line.x.bottom = element_line(arrow = arrow(type="closed")),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom")

p5
```

Use Matchmakers table, i.e. outcomes for each record. N.B. The actual table is not
present in the repo, one should write to the authors to obtain a copy.

```{r}
data.mm <- read_csv("matchmakers/01_05_2025_TCRvdb.csv")
data.mm <- rbind(tibble(cdr3 = data.mm$cdr3_alpha_aa,
                        v.segm = data.mm$TRAV,
                        padj = data.mm$padj),
                 tibble(cdr3 = data.mm$cdr3_beta_aa, 
                        v.segm = data.mm$TRBV,
                        padj = data.mm$padj)) |>
  group_by(cdr3) |>
  summarise(padj = min(padj))

data.slim |>
  filter(antigen.epitope %in% c("YLQPRTFLL", "GLCTLVAML"),
         complex.id != 0) |> # paired records only
  mutate(in.motif = ifelse(cdr3 %in% data.motifs$cdr3aa,
                           "TCRNET+", "orphan")) |>
  mutate(vdjdb.score = case_when(
    vdjdb.score == 1 ~ "Lo",
    vdjdb.score == 2 ~ "Med",
    vdjdb.score == 3 ~ "Hi",
    T ~ "?"
  ) |> factor(levels = c("?", "Lo", "Med", "Hi"))) |>
  separate_rows(reference.id, sep = ",") |>
  group_by(gene, v.segm, j.segm, cdr3) |>
  mutate(studies = length(unique(reference.id))) |>
  mutate(studies = ifelse(studies > 2, "3+", as.character(studies))) |>
  left_join(data.mm) |>
  mutate(matchmakers = ifelse(is.na(padj), "?", ifelse(padj < 1e-5, "MM+", "MM-"))) |>
  group_by(antigen.epitope, vdjdb.score, studies, in.motif, matchmakers) |>
  summarise(clonotypes = length(unique(paste(v.segm, j.segm, cdr3)))) -> data.alluvial.mm.2
```

```{r}
p6 <- ggplot(data = data.alluvial.mm.2,# |>
        # filter(vdjdb.score != "N/A"),
       aes(axis1 = studies,
           axis2 = antigen.epitope,
           axis4 = vdjdb.score,
           axis5 = in.motif,
           axis6 = matchmakers,
           y = clonotypes)) +
  scale_x_discrete("",
                   limits = c("Studies", "Antigen", "VDJdb score", "Motif", "Validation"), 
                   expand = c(.5, .05),
                   position = "bottom") +
  ylab("") +
  geom_alluvium(aes(fill = matchmakers), color = NA) +
  scale_fill_manual(guide = F, values = c("grey", "#de77ae", "#7fbc41")) +
  geom_stratum(color = "black", fill = "white", size = 0.2) +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            angle = 90, family="Helvetica light",
            cex = 3) +
  annotation_custom(img.vdjdb, xmin=-1.0, xmax=0.75, ymin=-Inf, ymax=Inf) +
  annotation_custom(img.mm, xmin=5.3, xmax=7, ymin=-Inf, ymax=Inf) +
  theme_classic() +
  theme(text = element_text(size=12, family="Helvetica light"),
        axis.line.y = element_blank(),
        axis.line.x.bottom = element_line(arrow = arrow(type="closed")),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom")

p6
```

Enrichment test for validated records in TCRNET+

```{r}
data.alluvial.mm.2 |>
  filter(matchmakers != "?") |>
  group_by(in.motif, matchmakers) |>
  summarise(clonotypes = sum(clonotypes)) -> data.alluvial.mm.s

data.alluvial.mm.s

dcast(data.alluvial.mm.s, in.motif ~ matchmakers) |>
  column_to_rownames("in.motif") |>
  fisher.test() -> odds

data.alluvial.mm.s$clonotypes[1] / data.alluvial.mm.s$clonotypes[2] / 
  (data.alluvial.mm.s$clonotypes[3] / data.alluvial.mm.s$clonotypes[4])
odds
odds$p.value
```

```{r fig.width=10.5, fig.height=13}
tmp1 <- ((p1 + p1a + (p2 / p3)) + plot_layout(widths = c(1.1, 1.00, 2.93)))
tmp2 <- tmp1 / p4 / p6
tmp2 +
  plot_layout(heights = c(2.0, 1.5, 3.5)) +
  plot_annotation(tag_levels = 'a') & 
  theme(plot.tag = element_text(size = 16, family = "Helvetica")) -> p_full

p_full

ggsave("figures/fig2.pdf", p_full, width = 10.5, height = 13)
```

### Considering the effect of V allele calling and pairing

Checking for ambiguity in V gene assignment

```{r}
data |>
  group_by(cdr3.beta) |>
  summarise(distinct.v.beta.vdjdb = length(unique(v.beta))) |>
  mutate(distinct.v.beta.vdjdb = ifelse(distinct.v.beta.vdjdb >= 2,
                                        "2+", distinct.v.beta.vdjdb)) |>
  group_by(distinct.v.beta.vdjdb) |>
  summarise(unique_cdr3b = n())

data |>
  group_by(cdr3.alpha) |>
  summarise(distinct.v.alpha.vdjdb = length(unique(v.alpha))) |>
  mutate(distinct.v.alpha.vdjdb = ifelse(distinct.v.alpha.vdjdb >= 2,
                                        "2+", distinct.v.alpha.vdjdb)) |>
  group_by(distinct.v.alpha.vdjdb) |>
  summarise(unique_cdr3a = n())

data.mm.1 <- read_csv("matchmakers/01_05_2025_TCRvdb.csv") |>
  mutate(matchmakers = ifelse(is.na(padj), "?", ifelse(padj < 1e-5, "MM+", "MM-"))) |>
  filter(matchmakers != "?") 

data.mm.1 |>
  group_by(cdr3.beta = cdr3_beta_aa) |>
  summarise(distinct.v.beta.mm = length(unique(TRBV_IMGT)) > 1,
            distinct.mm.outcome = length(unique(matchmakers)) > 1) |>
  group_by(distinct.v.beta.mm, distinct.mm.outcome) |>
  summarise(unique_cdr3b = n())

data.mm.1 |>
  group_by(cdr3.alpha = cdr3_alpha_aa) |>
  summarise(distinct.v.alpha.mm = length(unique(TRAV_IMGT)) > 1,
            distinct.mm.outcome = length(unique(matchmakers)) > 1) |>
  group_by(distinct.v.alpha.mm, distinct.mm.outcome) |>
  summarise(unique_cdr3a = n())
```

Distinct validation outcome due to pairing (CDR3 aa level)

```{r}
read_tsv("dump/vdjdb-2025-02-21/vdjdb_full.txt") |>
  filter(species == "HomoSapiens",
         antigen.epitope %in% c("YLQPRTFLL", "GLCTLVAML")) |>
  summarise(unpaired = sum(is.na(cdr3.beta) | is.na(cdr3.alpha)),
            paired = sum(!is.na(cdr3.beta) & !is.na(cdr3.alpha)))

data.mm.1 |>
  group_by(cdr3_beta_aa) |>
  summarise(distinct.alpha.pair = length(unique(cdr3_alpha_aa)) > 1,
            distinct.mm.outcome = length(unique(matchmakers)) > 1) |>
  group_by(distinct.alpha.pair, distinct.mm.outcome) |>
  summarise(unique_cdr3b = n())

data.mm.1 |>
  group_by(cdr3_alpha_aa) |>
  summarise(distinct.beta.pair = length(unique(cdr3_beta_aa)) > 1,
            distinct.mm.outcome = length(unique(matchmakers)) > 1) |>
  group_by(distinct.beta.pair, distinct.mm.outcome) |>
  summarise(unique_cdr3a = n())
```

```{r}
sessionInfo()
```